FL_abund_GAM

### 2020-09-10 analyses/BB_abund_GAM.Rmd
rearranged directories

---
title: "preference"
output: html_document
---
### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, year, date, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
```

### Load network data
```{r}
### BB-FL visitation data
net <- read_csv("../processed_data/network.csv") %>%
  dplyr::select(site, date, bb.sp, plant.sp)
```

### FL and BB abundance / diversity
```{r echo=FALSE, include=FALSE}
### Per-transect BB abundance: species-level
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
  ungroup %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  mutate(year = factor(year(date))) %>%
  left_join(site_data) %>%
  mutate(yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management)) %>%
  semi_join(net, by = c("site", "date")) %>%
  group_by(year, bb.sp) %>%
  mutate(bb.total.abund = sum(bb.abund))
```

GAM modeling of abundance per bb species
```{r}
### Call models
bb_abund.2010 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                            by = bb.sp,
                                                            bs= c("tp", "tp"), 
                                                            k=c(10, 10), m=1),
                        data = filter(bb_abund_sp, year == 2010 & bb.total.abund >= 5),
                        discrete = TRUE,
                        family = "poisson",
                        select = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(bb_abund.2010)
bb_abund.2010.sum <- summary(bb_abund.2010)
print(plot(bb_abund.2010, allTerms = TRUE), pages = 4)

bb_abund.2011 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                            by = bb.sp,
                                                            bs= c("tp", "tp"), 
                                                            k=c(10, 10), m=1),
                        data = filter(bb_abund_sp, year == 2011 & bb.total.abund >= 5),
                        discrete = TRUE,
                        family = "poisson",
                        select = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(bb_abund.2011)
bb_abund.2011.sum <- summary(bb_abund.2011)
print(plot(bb_abund.2011 , allTerms = TRUE), pages = 4)

bb_abund.2012 <- bam(bb.abund ~ s(bb.sp, bs = "re") + management + slope.est + te(yday, elev.mean, 
                                                            by = bb.sp,
                                                            bs= c("tp", "tp"), 
                                                            k=c(10, 10), m=1),
                        data = filter(bb_abund_sp, year == 2012 & bb.total.abund >= 5),
                        discrete = TRUE,
                        family = "poisson",
                        select = TRUE,
                        method = "fREML") %>% getViz()

check.gamViz(bb_abund.2012)
bb_abund.2012.sum <- summary(bb_abund.2012)
print(plot(bb_abund.2012 , allTerms = TRUE), pages = 4)
plot(bb_abund.2012)

```

```{r}
bb_abund.2012.pred <- predict(bb_abund.2012)

ggplot(bb_abund.2012.pred, aes(yday, elev.mean, fit)) +
  geom_raster(aes(fill = exp(fit))) +
  #geom_contour(aes(z = fit), color = "black") +
  geom_point(data = filter(fl_abund, year == 2012), 
             aes(yday, elev.mean, size = fl.abund),
             color = "white", alpha = 0.25) +
  scale_fill_viridis_c()
```

>>>>

---
title: "BB_abund_GAM"
output: html_document
---
### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>% # make year a factor
  dplyr::select(site, year, date, gdd.cum) %>% # drop unused columns
  semi_join(read_csv("../processed_data/network.csv"), by = "site") # drop sites that were not samples
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")), # rename in English
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
```

### Load network data
```{r}
### BB-FL visitation data
net <- read_csv("../processed_data/network.csv") %>%
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns
```

### BB abundance / diversity
```{r echo=FALSE, include=FALSE}
### Per-transect BB abundance: species-level
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  group_by(site, date, bb.sp) %>% # 
  summarize(bb.abund = n()) %>% # get per transect bb abundance
  ungroup() %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  semi_join(net, by = c("site", "date")) %>% # drop site-dates that did not exist in the data set
  left_join(site_data) %>% # add site data
  mutate(year = factor(year(date)),
         yday = yday(date),
         bb.sp = factor(bb.sp),
         management = factor(management),
         site = factor(site),
         transect = factor(transect)) %>%
  group_by(year, bb.sp) %>%
  mutate(bb.total.abund = sum(bb.abund))
```

## Global GAM: this is more elegant than breaking it down by year, but my concern is that the model will be totally dominated by the 2012 data, since I have so many more samples that year.
```{r}
bb_abund.global <- bam(bb.abund ~
                         s(bb.sp, bs = "re") +
                         s(year, bs = "re") +
                         s(site, bs = "re") + ### I think I need to include site as a random effect because if I don't the pseudoreplication of elevation is pretty extreme, especially in 2012 where I have lots of repeated measures. I am not, for example, measuring floral abundance at 1000 m 20 independent times; I'm measuring it 20 times at the same spot. This does, of  course, create identifiability issues insofar as site and elevation are concurved. But I think I need to accept these problems.
                         te(yday, elev.mean, 
                                   by = bb.sp,
                                   #id = 0,
                                   bs= c("gp", "tp"), 
                                   k=c(8, 8), 
                                   m=1) +
                         te(yday, elev.mean, 
                                   by = year,
                                   #id = 0,
                                   bs= c("gp", "tp"), 
                                   k=c(4, 4), 
                                   m=2),
                       data = filter(bb_abund_sp, bb.total.abund >= 5),
                       discrete = TRUE,
                       family = "tw",
                       select = TRUE,
                       method = "fREML") %>% getViz(nsim = 500)

bb_abund.global2 <- gam(bb.abund ~
                s(bb.sp, bs = "re") +
                s(year, bs = "re") +
                s(site, bs = "re") +
                te(yday, elev.mean, 
                          by = bb.sp,
                          #id = 0,
                          bs= c("gp", "tp"), 
                          k=c(6, 6), 
                          m=1) +
                te(yday, elev.mean, 
                          by = year,
                          #id = 0,
                          bs= c("gp", "tp"), 
                          k=c(4, 4), 
                          m=2),
              data = filter(bb_abund_sp, bb.total.abund >= 5 & bb.sp %in% c("bss", "pasc", "prat", "pyre")),
              discrete = FALSE,
              family = ziP(),
              select = TRUE,
              method = "REML") %>% getViz(nsim = 500)

ziptest <- gam(bb.abund ~
                t2(yday, elev.mean, 
                          by = bb.sp,
                          #id = 0,
                          bs= c("gp", "tp"), 
                          k=c(3, 3), 
                          m=1),
              data = filter(bb_abund_sp, bb.total.abund >= 5),
              #discrete = FALSE,
              family = ziP(),
              select = TRUE,
              method = "REML") %>% getViz(nsim = 50)

check.gamViz(bb_abund.global)
check.gamViz(bb_abund.global2)
check.gamViz(ziptest)

bb_abund.global.sum <- summary(bb_abund.global)
bb_abund.global2.sum <- summary(bb_abund.global2)

AIC(bb_abund.global, bb_abund.global2)

print(plot(bb_abund.global, allTerms = TRUE), pages = 3)
print(plot(bb_abund.global2, allTerms = TRUE), pages = 3)

bb_abund.global.ck <- check2D(bb_abund.global, x1 = "yday", x2 = "elev.mean")
bb_abund.global.ck + l_gridCheck2D(gridFun = mean)
```

## Annual GAMs
### My argument for not including site as a random effect is that it would create identifiability issues with the elevation variable that I am actally interested in. This will probably cause the p-values to be underestimated, maybe to the point of being uninformative. But I don't think I care. My goal is not to establish inferentially that time and elevation affect BB abundance; I want to say how they do. And I don't want to obscure the "how" for the sake of inferential scruples.  
```{r}
### 2010
bb_abund.2010 <- bam(bb.abund ~ 
                       s(bb.sp, bs = "re") + 
                       te(yday, elev.mean, 
                          by = bb.sp,
                          bs= c("gp", "cr"), 
                          k=c(8, 8), 
                          m=2),
                     data = filter(bb_abund_sp, year == 2010 & bb.total.abund >= 5),
                     discrete = TRUE,
                     family = "tw",
                     select = TRUE,
                     method = "fREML") %>% getViz(nsim = 500)

check.gamViz(bb_abund.2010)
bb_abund.2010.sum <- summary(bb_abund.2010)
print(plot(bb_abund.2010, allTerms = TRUE), pages = 1)
bb_abund.2010.ck <- check2D(bb_abund.2010, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

### 2011
bb_abund.2011 <- bam(bb.abund ~ 
                       s(bb.sp, bs = "re") + 
                       te(yday, elev.mean, 
                          by = bb.sp,
                          bs= c("gp", "cr"), 
                          k=c(8, 8), 
                          m=2),
                     data = filter(bb_abund_sp, year == 2011 & bb.total.abund >= 5),
                     discrete = TRUE,
                     family = "tw",
                     select = TRUE,
                     method = "fREML") %>% getViz(nsim = 500)

check.gamViz(bb_abund.2011)
bb_abund.2011.sum <- summary(bb_abund.2011)
print(plot(bb_abund.2011, allTerms = TRUE), pages = 1)
bb_abund.2011.ck <- check2D(bb_abund.2011, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals

### 2012
bb_abund.2012 <- bam(bb.abund ~ 
                       s(bb.sp, bs = "re") + 
                       te(yday, elev.mean, 
                          by = bb.sp,
                          bs= c("gp", "cr"), 
                          k=c(8, 8), 
                          m=2),
                     data = filter(bb_abund_sp, year == 2012 & bb.total.abund >= 5),
                     discrete = TRUE,
                     family = "tw",
                     select = TRUE,
                     method = "fREML") %>% getViz(nsim = 500)

check.gamViz(bb_abund.2012)
bb_abund.2012.sum <- summary(bb_abund.2012)
print(plot(bb_abund.2012, allTerms = TRUE), pages = 1)
bb_abund.2012.ck <- check2D(bb_abund.2012, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
```

### Tabulate plots
```{r}
bss2010 <- plot(sm(bb_abund.2010, 2)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
bss2011 <- plot(sm(bb_abund.2011, 2)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)
bss2012 <- plot(sm(bb_abund.2012, 2)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. terrestris/lucorum")), x = NULL, y = NULL)

gers2010 <- plot(sm(bb_abund.2010, 3)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)
gers2011 <- ggplot() #too rare
gers2012 <- plot(sm(bb_abund.2012, 3)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. gerstaeckeri")), x = NULL, y = NULL)

hort2010 <- plot(sm(bb_abund.2010, 4)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
hort2011 <- plot(sm(bb_abund.2011, 3)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)
hort2012 <- plot(sm(bb_abund.2012, 4)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. hortorum")), x = NULL, y = NULL)

hypn2010 <- plot(sm(bb_abund.2010, 5)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
hypn2011 <- plot(sm(bb_abund.2011, 4)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. hypnorum")), x = NULL, y = NULL)
hypn2012 <- ggplot()  #too rare

jone2010 <- plot(sm(bb_abund.2010, 6)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
jone2011 <- plot(sm(bb_abund.2011, 5)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)
jone2012 <- plot(sm(bb_abund.2012, 5)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. jonellus")), x = NULL, y = NULL)

lapi2010 <- ggplot() # too rare 
lapi2011 <- plot(sm(bb_abund.2011, 6)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)
lapi2012 <- plot(sm(bb_abund.2012, 6)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. lapidarius")), x = NULL, y = NULL)

mend2010 <- plot(sm(bb_abund.2010, 7)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
mend2011 <- plot(sm(bb_abund.2011, 7)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)
mend2012 <- plot(sm(bb_abund.2012, 7)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. mendax")), x = NULL, y = NULL)

mont2010 <- plot(sm(bb_abund.2010, 8)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
mont2011 <- plot(sm(bb_abund.2011, 8)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE)+
  l_points() +
  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)
mont2012 <- plot(sm(bb_abund.2012, 8)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. monticola")), x = NULL, y = NULL)

muci2010 <- plot(sm(bb_abund.2010, 9)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
muci2011 <- plot(sm(bb_abund.2011, 9)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)
muci2012 <- plot(sm(bb_abund.2012, 9)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. mucidis")), x = NULL, y = NULL)

pasc2010 <- plot(sm(bb_abund.2010, 10)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
pasc2011 <- plot(sm(bb_abund.2011, 10)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)
pasc2012 <- plot(sm(bb_abund.2012, 10)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. pascuorum")), x = NULL, y = NULL)

prat2010 <- plot(sm(bb_abund.2010, 11)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
prat2011 <- plot(sm(bb_abund.2011, 11)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)
prat2012 <- plot(sm(bb_abund.2012, 11)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. pratorum")), x = NULL, y = NULL)

psit2010 <- plot(sm(bb_abund.2010, 12)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
psit2011 <- plot(sm(bb_abund.2011, 12)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)
psit2012 <- plot(sm(bb_abund.2012, 12)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. psithyrus")), x = NULL, y = NULL)

pyre2010 <- plot(sm(bb_abund.2010, 13)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
pyre2011 <- plot(sm(bb_abund.2011, 13)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)
pyre2012 <- plot(sm(bb_abund.2012, 13)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. pyrenaeus")), x = NULL, y = NULL)

soro2010 <- plot(sm(bb_abund.2010, 14)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
soro2011 <- plot(sm(bb_abund.2011, 14)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)
soro2012 <- plot(sm(bb_abund.2012, 14)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. soroensis")), x = NULL, y = NULL)

wurf2010 <- plot(sm(bb_abund.2010, 15)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
wurf2011 <- plot(sm(bb_abund.2011, 15)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  guides(fill = FALSE) +
  l_points() +
  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)
wurf2012 <- plot(sm(bb_abund.2012, 15)) +
  l_fitRaster() +
  l_fitContour(size = 0.1) +
  l_points() +
  guides(fill = FALSE) +
  labs(title = expression(italic("B. wurflenii")), x = NULL, y = NULL)

pdf("./output/BB_abund_major.pdf", height = 11, width = 8.5)
major_sp <- gridPrint(bss2010, bss2011, bss2012, 
                      pasc2010, pasc2011, pasc2012,
                      prat2010, prat2011, prat2012,
                      psit2010, psit2011, psit2012,
                      soro2010, soro2011, soro2012,
                      wurf2010, wurf2011, wurf2012, ncol = 3)
dev.off()

pdf("./output/BB_abund_highelev.pdf", height = 5.5, width = 8.5)
high_elev_sp <- gridPrint(pyre2010, pyre2011, pyre2012, 
                          mend2010, mend2011, mend2012,
                          mont2010, mont2011, mont2012, ncol = 3)
dev.off()

pdf("./output/BB_abund_minor.pdf", height = 11, width = 8.5)
minor_sp <- gridPrint(gers2010, gers2011, gers2012,
                      hort2010, hort2011, hort2012,
                      hypn2010, hypn2011, hypn2012,
                      jone2010, jone2011, jone2012,
                      lapi2010, lapi2011, lapi2012,
                      muci2010, muci2011, muci2012, ncol = 3)
dev.off()
```

## Bumble bee species richness
```{r}
bb_rich <- bb_abund_sp %>%
  filter(bb.abund > 0) %>%
  group_by(site, date, yday, year, elev.mean, management) %>%
  summarize(bb.rich = length(unique(bb.sp)))
```

## Global species richness GAM (no need to separate the years, meiner Meinung nach)
```{r}
### 2010
bb_rich.global <- bam(bb.rich ~ 
                        te(yday, elev.mean, 
                           by = year,
                           bs= c("gp", "cr"), 
                           k=c(6, 6), 
                           m=2),
                      data = filter(bb_rich),
                      discrete = TRUE,
                      family = "gaussian",
                      select = TRUE,
                      method = "fREML") %>% getViz(nsim = 500)

check.gamViz(bb_rich.global)
bb_rich.global.sum <- summary(bb_rich.global)
print(plot(bb_rich.global, allTerms = TRUE), pages = 1)
bb_rich.global.ck <- check2D(bb_rich.global, x1 = "yday", x2 = "elev.mean") # See if there is any spatiotemporal pattern to the residuals
```

### 2020-09-10 analyses/FL_abund_GAM.Rmd
rearranged directories

---
title: "preference"
output: html_document
---
### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>%
  dplyr::select(site, year, date, gdd.cum) %>%
  semi_join(read_csv("../processed_data/network.csv"), by = "site")
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  #dplyr::select(-c(slope.calc, slope.est, elev.min, elev.max, temp.mean)) %>% # drop these variables
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")),
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
```

### Load network and survey data
```{r}
### BB-FL visitation data
net <- read_csv("../processed_data/network.csv") %>%
  dplyr::select(site, date, bb.sp, plant.sp)

### Floral survey data
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only plants visited (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)

survey_dups <- survey %>%
  group_by(site, date, plant.sp) %>%
  mutate(dups = n()) %>%
  filter(dups > 1)

dup_species <- survey_dups %>%
  group_by(plant.sp) %>%
  mutate(dups = n()/2) %>%
  group_by(site, date, plant.sp, dups) %>%
  mutate(different = length(unique(flower.cover)) -1) %>%
  group_by(site, date, plant.sp, dups) %>%
  summarize(different = mean(different)) %>%
  group_by(plant.sp, dups) %>%
  summarize(different.dups = sum(different)) %>% 
  arrange(-dups)

@@

### FL and BB abundance / diversity
```{r echo=FALSE, include=FALSE}
### Per-transect BB abundance: species-level
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  semi_join(shared_transects) %>% # I only want BB data with corresponding floral data
  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
  ungroup %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  mutate(year = factor(year(date))) %>%
  semi_join(net, by = c("site", "date")) # only include site-dates that actually occurred

@@

### 2010
weighted_fl_abundance.2010 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(10, 10), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2010),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2010)
weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.2010.ck + l_gridCheck2D(gridFun = mean)

>>>>

---
title: "FL_abund_GAM"
output: html_document
---
### Prepare environment
```{r setup, include=FALSE, echo=FALSE}
library(mgcv)
library(mgcViz)
library(tidyverse)
library(lubridate)
```

### Climate data
```{r, include=FALSE, echo=FALSE}
gdd <- read_csv("../processed_data/climate.csv") %>%
  mutate(year = factor(year)) %>% # make year a factor
  dplyr::select(site, year, date, gdd.cum) %>% # drop unused columns
  semi_join(read_csv("../processed_data/network.csv"), by = "site") # drop sites that were not samples
```

### Load site data
```{r, include=FALSE, echo=FALSE}
site_data <- read_csv("../processed_data/site_data.csv") %>%
  mutate(elev.class = factor(elev.class, levels = c("oben", "mitte", "unten")), # rename in English
         elev.class2 = factor(elev.class2, levels = c("high", "mid", "low")))
```

### Load network and survey data
```{r}
### BB-FL visitation data
net <- read_csv("../processed_data/network.csv") %>%
  dplyr::select(site, date, bb.sp, plant.sp) # drop unused columns

### Floral survey data
survey <- read_csv("../processed_data/floral_survey.csv") %>%
  semi_join(net, by = "plant.sp") %>% # Consider only visited plant species (anytime anywhere, not specifically for a given site*date)
  dplyr::select(site, date, plant.sp, flower.cover)

### Identify duplicate observations; these may be errors or intentionally duplicated (see correspondence with Kalli)
survey_dups <- survey %>%
  group_by(site, date, plant.sp) %>%
  mutate(dups = n()) %>%
  filter(dups > 1)

@@

### FL and BB abundance / diversity
```{r echo=FALSE, include=FALSE}
### Per-transect BB abundance: species-level
bb_abund_sp <- net %>%
  dplyr::select(site, date, bb.sp) %>%
  semi_join(shared_transects) %>% # I only want BB data with corresponding floral data
  group_by(site, date, bb.sp) %>% # group by bb.sp*transect (and other variables we want to retain)
  summarize(bb.abund = n()) %>% # get bb.sp*transect abundance
  ungroup %>%
  complete(site, date, bb.sp, fill = list(bb.abund = 0)) %>% # complete implicit absences
  mutate(year = factor(year(date))) %>%
  semi_join(shared_transects, by = c("site", "date")) # only include site-dates that actually occurred

@@

### 2010
weighted_fl_abundance.2010 <- bam(log(fl.abund.wt + 0.001) ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(10, 10), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2010),
                                    discrete = TRUE,
                                    family = "gaussian",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

weighted_fl_abundance.2010b <- bam(fl.abund.wt + 0.001 ~
                                      s(bb.sp, bs = "re") +
                                      #s(management, bs = "re") +
                                      #s(site, bs = "re") +
                                      te(yday, elev.mean, 
                                                by = bb.sp,
                                                #id = 0,
                                                bs= c("gp", "tp"), 
                                                k=c(15, 15), 
                                                m=2),
                                    data = filter(fl_abund_wt, year == 2010),
                                    discrete = TRUE,
                                    family = "tw",
                                    select = TRUE,
                                    method = "fREML") %>% getViz(nsim = 500)

check.gamViz(weighted_fl_abundance.2010)
check.gamViz(weighted_fl_abundance.2010b)

weighted_fl_abundance.2010.sum <- summary(weighted_fl_abundance.2010)
print(plot(weighted_fl_abundance.2010, allTerms = TRUE), pages = 4)
print(plot(weighted_fl_abundance.2010b, allTerms = TRUE), pages = 4)

plot(sm(weighted_fl_abundance.2010, 2))
plot(sm(weighted_fl_abundance.2010b, 2))

weighted_fl_abundance.2010.ck <- check2D(weighted_fl_abundance.2010, 
                                         x1 = "yday", x2 = "elev.mean")
weighted_fl_abundance.2010b.ck <- check2D(weighted_fl_abundance.2010b, 
                                         x1 = "yday", x2 = "elev.mean")

weighted_fl_abundance.2010.ck + l_gridCheck2D(gridFun = mean)
